Designing Qusecs Optimally: Quasi-uniform or Self-similar Layered Materials as Recursively Decomposed Solutions of Geometric Constraint Systems

### 2015-03-25 doc/GMP2015/revision_2/appendix_defs.tex
Started revision 2 directory. Got rid of \cutout, \addin, and \notetoreviewers macros.

\notetoreviewer{The following section was moved in from the Preliminaries section.}

@@

In this paper, we will consider 3 types of constraint systems: bar-joint as well as \dfn{body-hyperpin} \addin{(defined formally in Section \ref{sec:bodypin})} and \dfn{pinned line-incidence} \addin{(defined formally in Section \ref{sec:pinnedline})}. In all of these cases, a Cartesian \dfn{realization} or \dfn{solution} $G(p)$ of $(G,\delta)$ is an assignment of coordinates or Euclidean transformations (poses), $p: V \rightarrow \mathbb{R}^2$ or $\mathbb{R}^3$, to the vertices of $G$ satisfying the constraints with parameters $\delta$, modulo orientation preserving isometries (Euclidean rigid body motions for distance constraint).

@@

\cutout{In this paper, a qusecs is any independent geometric constraint system of one of the 3 types mentioned above.} \notetoreviewer{This now only in the Preliminaries section.}

>>>>

@@

In this paper, we will consider 3 types of constraint systems: bar-joint as well as \dfn{body-hyperpin} (defined formally in Section \ref{sec:bodypin}) and \dfn{pinned line-incidence} (defined formally in Section \ref{sec:pinnedline}). In all of these cases, a Cartesian \dfn{realization} or \dfn{solution} $G(p)$ of $(G,\delta)$ is an assignment of coordinates or Euclidean transformations (poses), $p: V \rightarrow \mathbb{R}^2$ or $\mathbb{R}^3$, to the vertices of $G$ satisfying the constraints with parameters $\delta$, modulo orientation preserving isometries (Euclidean rigid body motions for distance constraint).

@@

### 2015-03-25 doc/GMP2015/revision_2/appendix_proofs.tex
Started revision 2 directory. Got rid of \cutout, \addin, and \notetoreviewers macros.

For (4), because subgraphs of a isostatic graph can only be trivial, under\addin{constrained}, or isostatic, all cases have already been exhausted.
\end{proof}

@@

\subsubsection{Proof of Observation \ref{obs:algebraic_completion}}
\begin{proof}
    An isostatic graph has 3 parameters that define it (in terms of coming up with a realizaiton). These are the Euclidean motions. A 1-dof graph can be seen as essentially having 4 parameters: the 3 Euclidean motions and whatever degree of freedom it has. A 2-dof has 5 parameters, and so on. These paramters denote the algebraic complexity of giving a realization for the graphs.

@@

    For the 2-dof case, there are more cases to consider. If we pin the same node twice as above, we have $np(v) = 5nc_2(v)+3$ for any ancestor $v \neq b$ and $np(b) = 5nc_2(b)$. If we pin a node $v$ and one of its ancestors $v'$, then any nodes between $v'$ and $v$ will be 1-dof, any nodes above $v'$ will be isostatic, and nodes below $v$ will be 2-dof. Note that solving $v'$ will result in solving $v$. Then, we need to consider nodes above and includiing $v'$ in our complexity. $np(v') = 5nc_2(v') + 4$ and $np(a) = 3 + 5nc_2(a)$ for $a$ an ancestor of $v'$. The only remaining case will be if we pin two nodes that are not descendent/ancestor. the only change from the previous case will be that for the lowest common ancestor of the nodes $v'$, $np(v') = 2*4+5nc_2(v')$. For any ancestor $a$ of $v'$, we still have $np(a) = 3 + 5nc_2(a)$.

>>>>

For (4), because subgraphs of a isostatic graph can only be trivial, underconstrained, or isostatic, all cases have already been exhausted.
\end{proof}

@@

\subsubsection{Proof of Observation \ref{obs:algebraic_completion}}
\begin{proof}
    An isostatic graph has 3 parameters that define it (in terms of coming up with a realization). These are the Euclidean motions. A 1-dof graph can be seen as essentially having 4 parameters: the 3 Euclidean motions and whatever degree of freedom it has. A 2-dof has 5 parameters, and so on. These parameters denote the algebraic complexity of giving a realization for the graphs.

@@

    For the 2-dof case, there are more cases to consider. If we pin the same node twice as above, we have $np(v) = 5nc_2(v)+3$ for any ancestor $v \neq b$ and $np(b) = 5nc_2(b)$. If we pin a node $v$ and one of its ancestors $v'$, then any nodes between $v'$ and $v$ will be 1-dof, any nodes above $v'$ will be isostatic, and nodes below $v$ will be 2-dof. Note that solving $v'$ will result in solving $v$. Then, we need to consider nodes above and including $v'$ in our complexity. $np(v') = 5nc_2(v') + 4$ and $np(a) = 3 + 5nc_2(a)$ for $a$ an ancestor of $v'$. The only remaining case will be if we pin two nodes that are not descendant/ancestor. the only change from the previous case will be that for the lowest common ancestor of the nodes $v'$, $np(v') = 2*4+5nc_2(v')$. For any ancestor $a$ of $v'$, we still have $np(a) = 3 + 5nc_2(a)$.

### 2015-03-25 doc/GMP2015/revision_2/s01_introduction.tex
Started revision 2 directory. Got rid of \cutout, \addin, and \notetoreviewers macros.

\section{Response to reviewers}

\cutout{This means} the text was removed.
\addin{This means} the text was added.
\notetoreviewer{This means} it is a note to the reviewer, and would not appear in real document.
% \movedfrom{This means} the text was moved from another part of the document.

\subsection{Reviewer 1}
No suggestions made.

\subsection{Reviewer 2}
% \noindent
% Changes made:
\begin{itemize}
  \item We added a new example in figures to illustrate definitions/proofs/comparisons. And we added a new conjecture relating this method to the Modified Frontier Algorithm, the most recent and best current algorithm.
  \item Abstract was shortened and clarified.
  \item Basic definitions from Section \ref{sec:prelim} were moved to the new \ref{sec:appendix:defs}. Definition of qusecs was made more prominent, left in Section \ref{sec:prelim}. Slight modification to definition of DR-plan.
  \item Subsections \ref{sec:cont} and \ref{sec:intro:prevwork} were flipped. This gives readers an idea of the contributions before they read about the current state of the field.
  \item Section \ref{sec:DRP} had some proofs moved back in from the appendix. Section \ref{sec:recomb} had a proof moved back in from the appendix. Section \ref{sec:conclusion} had future work moved back in from the appendix.
  \item Numerous grammatical, syntactical, and spelling errors were fixed. Made hyphen usage consistent.
  \item Images were properly aligned and cropped.
\end{itemize}

\subsection{Reviewer 3}
We emailed the co-chairs, and they replied that it would fit in the conference, leaving it at our discretion to withdraw. We have decided to continue.

@@

Some properties of natural and engineered materials can be analyzed by treating them as two dimensional (2D) layers. As illustrated by the examples below, the structure within each layer is often self-similar~\cite{2012arXiv1204.6389G} spanning multiple scales; generally aperiodic and quasi-uniform within any one scale; and consists of a few repeated motifs appearing in disordered arrangements. Each layer is not necessarily planar, i.e., it consists of multiple, inter-constraining planar (genus 0) monolayers. Furthermore, a layer is often  either
\vemph{isostatic or underconstrained}\cutout{,} \addin{(}\vemph{not self-stressed}\addin{,}
\cutout{(}\seedefsd).
\cutout{These properties are consistent with a self-assembled 2D structure that minimizes mass, optimally distributes external stresses and itself participates in the assembly of diverse and multifunctional, larger 2D structures.}
\addin{A variety of quasi-uniform (aperiodic) and self-similar, layered material structures are a natural consequence of design objectives such as rigidity or flexibility, minimizing mass, optimally distributing external stresses and participating in the assembly of diverse and multifunctional, larger  structures.}

@@

\notetoreviewer{The position of this section and following section were swapped.}

The contributions of this paper are the following.
\begin{itemize}
  \item \addin{In Section \ref{sec:prelim}, we give definitions and give the new notion of qusecs and the canonical/optimal DR-plan. This is relevant to Examples \ref{materialexample1} and \ref{materialexample2}, where we model them as bar-joint systems and discuss achieving isostaticity, distribution of stresses in self-similar, and other important concepts.}

@@

  \item Also in Section \ref{sec:DRP}, we give an efficient (\candrpcomplexity) algorithm to find a canonical (and hence optimal) DR-plan for all 3 types of 2D qusecs mentioned above (Sections \ref{sec:DRP}, \ref{sec:bodypin}, and \ref{sec:pinnedline}). The canonical DR-plan elucidates the essence of the NP-hardness of finding optimal DR-plans for over-constrained systems. Furthermore, our optimal/canonical DR-plan satisfies desirable properties such as the previously studied Cluster Minimality \cite{hoffman2001decompositionI} \addin{(see Figure~\ref{fig:demo_graph:clustmindrp})}.

@@

  \item In Section \ref{sec:bodypin} and \ref{sec:pinnedline}, we briefly describe applications of the above techniques to modeling, analyzing, and designing specific properties in 2D material layers~\cite{Jackson2008bodypin}.
  \cutout{For Examples 1 and 2 (achieving isostaticity, distribution of stresses in self-similar, bar-joint systems); Example 3 (canonical and optimal DR-plans for pinned line incidence systems [X]) and Examples 4 and 5 (boundary-conditions for achieving various desired properties of body-hyperpin systems).}
  \addin{For Examples \ref{materialexample4} and \ref{materialexample5}, we discuss boundary-conditions for achieving various desired properties of body-hyperpin systems. For Example \ref{materialexample3}, we discuss canonical and optimal DR-plans for pinned line incidence systems \cite{sitharam2014incidence}).}

@@

We now briefly survey existing techniques for studying 2D qusecs,  many of which are \dfn{bar-joint} systems (Examples 1\cutout{,} \addin{and} 2 above, see Sections \ref{sec:prelim}, \ref{sec:DRP}, \addin{and} \ref{sec:recomb}), \dfn{body-hyperpin} systems (Example 4\cutout{,} \addin{and} 5, see Section \ref{sec:bodypin}) or \dfn{pinned-line incidence} systems (Example 3, see Section \ref{sec:pinnedline}). The limitations of these techniques directly motivate the contributions of this paper.

@@

\medskip\noindent
\header{(ii) Optimal Recursive Decomposition (DR-planning).}
Recursive decomposition of geometric constraint systems has been formalized \cite{hoffman2001decompositionI,hoffman2001decompositionII} and well-studied \cite{jermann2006decomposition,sitharam2005combinatorial} as the \dfn {Decomposition-Recombination (DR-) planning} problem \seedefsprelim. For the abovementioned classes of 2D qusecs, generic rigidity is a combinatorial property and hence each level of the decomposition should, in principle, be achievable by a graph-based algorithm as in (1), without involving geometric information in the constraint system. Since many such decompositions can exist for a given constraint system, criteria defining desirable or optimal DR-plans and DR-planning algorithms were given in \cite{hoffman2001decompositionI}. An \dfn{optimal DR-plan} is one that minimizes the \dfn{size} \seedefsprelim, i.e., the maximum number of child subsystems of any parent system. Being exponential \addin{in} the size, the complexity of solving the parent constraint system \cutout{from the solutions of the of the child systems, is overwhelmingly dominated by it} \addin{is overwhelmingly dominated by the complexity of solving the child systems}.

@@

\medskip\noindent
\header{(iii) DR-plans for special classes and with other
criteria.}
For a special class of 2D qusecs, namely \dfn{tree-decomposable} systems  \cite{fudos1997graph,owen1991algebraic,joan-arinyo2004revisiting}  common in computer aided mechanical design, (which includes ruler-and-compass and Henneberg-I constructible systems), all DR-plans turn out to be optimal. This satisfies the so-called \dfn {Church-Rosser} criterion, leading to highly efficient DR-planning algorithms. For general 2D qusecs, alternate criteria were suggested such as \dfn{cluster minimality} requiring parent systems to be composed of a minimal set of at least 2 rigid proper subsystems (i.e., no proper subset forms a rigid system); and \dfn{proper maximality}, requiring child subsystems to be maximal rigid proper subsystems of the parent system. \seedefsc.

@@

\medskip\noindent
\header{(iv) Optimal Recombination and Solution Space Navigation.}
For the entire DR-plan, finding all desired solutions is barely tractable even if recombination of solved subsystems is easy for each indecomposable parent system in the DR-plan. This is because even for the simplest, highly decomposable systems with small DR-plans, the problem of finding even a single solution to the input system at the root of the DR-plan is NP-hard  \cite{saxe1979embeddability} and there is a combinatorial explosion of solutions \cite{borcea2004number}. Typically, however, the desired solution has a given orientation type, in which case, the crux of the difficulty is concentrated in the algebraic complexity of (re)combining child system solutions\cutout{,} to give a solution to the parent system at any given level of the DR-plan. For fairly general 3D constraint systems, there are algorithms with formal guarantees that uncover underlying matroids to combinatorially obtain an optimal parameterization to minimize the algebraic complexity of the indecomposable parent (sub)systems that occur in the DR-plan \cite{sitharam2010optimized,sitharam2006well,sitharam2010reconciling}, provided the DR-plan meets some of the abovementioned criteria.

@@

\medskip\noindent
\header{(v) Configuration Spaces of Underconstrained Systems.}
For underconstrained 2D bar-joint and body-hyperpin qusecs obtained from various subclasses of tree-decomposable systems, algorithms have been developed to complete them into isostatic systems \cite{joan-arinyo2003transforming,sitharam2005combinatorial,gao2006ctree,sitharam2010convex} and to find paths within the connected components \cite{sitharam2011cayleyI,hidalgo2011reachability} of standard Cartesian configuration spaces. Most of the algorithms with formal guarantees leverage Cayley configuration space theory \cite{sitharam2010convex,sitharam2011cayleyI,sitharam2011cayleyII} to characterize subclasses of graphs and additional constraints that control combinatorial explosion, and provide faithful bijective representation of connected components and paths. These algorithms have decreasing efficiency as the subclass of systems gets bigger, with highest efficiency for underlying partial 2-tree graphs (alternately called\cutout{,} tree-width 2, series-parallel, \addin{and} $K_4$ minor avoiding), moderate efficiency for 1 degree-of-freedom (dof) graphs with low Cayley complexity (which include common linkages such as the Strandbeest, Limacon and Cardioid), and decreased efficiency for general 1-dof tree-decomposable graphs. While software suites exist  \cite{keycurriculum1995geometer,porta2014open,siemens1999d,todd2007geometry}, no such formal algorithms and guarantees are known for non-tree-decomposable systems.

@@

\cutout{
\noindent
\textbf{Note:} All proofs appear in the Appendix.
% \sidenote{All proofs appear in X.}
}

>>>>

% \section{Response to reviewers}

% \cutout{This means} the text was removed.
% \addin{This means} the text was added.
% \notetoreviewer{This means} it is a note to the reviewer, and would not appear in real document.
% % \movedfrom{This means} the text was moved from another part of the document.

% \subsection{Reviewer 1}
% No suggestions made.

% \subsection{Reviewer 2}
% % \noindent
% % Changes made:
% \begin{itemize}
%   \item We added a new example in figures to illustrate definitions/proofs/comparisons. And we added a new conjecture relating this method to the Modified Frontier Algorithm, the most recent and best current algorithm.
%   \item Abstract was shortened and clarified.
%   \item Basic definitions from Section \ref{sec:prelim} were moved to the new \ref{sec:appendix:defs}. Definition of qusecs was made more prominent, left in Section \ref{sec:prelim}. Slight modification to definition of DR-plan.
%   \item Subsections \ref{sec:cont} and \ref{sec:intro:prevwork} were flipped. This gives readers an idea of the contributions before they read about the current state of the field.
%   \item Section \ref{sec:DRP} had some proofs moved back in from the appendix. Section \ref{sec:recomb} had a proof moved back in from the appendix. Section \ref{sec:conclusion} had future work moved back in from the appendix.
%   \item Numerous grammatical, syntactical, and spelling errors were fixed. Made hyphen usage consistent.
%   \item Images were properly aligned and cropped.
% \end{itemize}

% \subsection{Reviewer 3}
% We emailed the co-chairs, and they replied that it would fit in the conference, leaving it at our discretion to withdraw. We have decided to continue.

@@

Some properties of natural and engineered materials can be analyzed by treating them as two dimensional (2D) layers. As illustrated by the examples below, the structure within each layer is often self-similar~\cite{2012arXiv1204.6389G} spanning multiple scales; generally aperiodic and quasi-uniform within any one scale; and consists of a few repeated motifs appearing in disordered arrangements. Each layer is not necessarily planar, i.e., it consists of multiple, inter-constraining planar (genus 0) monolayers. Furthermore, a layer is often  either
\vemph{isostatic or underconstrained} (\vemph{not self-stressed},\seedefsd). A variety of quasi-uniform (aperiodic) and self-similar, layered material structures are a natural consequence of design objectives such as rigidity or flexibility, minimizing mass, optimally distributing external stresses and participating in the assembly of diverse and multifunctional, larger  structures.

@@

The contributions of this paper are the following.
\begin{itemize}
  \item In Section \ref{sec:prelim}, we give definitions and give the new notion of qusecs and the canonical/optimal DR-plan. This is relevant to Examples \ref{materialexample1} and \ref{materialexample2}, where we model them as bar-joint systems and discuss achieving isostaticity, distribution of stresses in self-similar, and other important concepts.

@@

  \item Also in Section \ref{sec:DRP}, we give an efficient (\candrpcomplexity) algorithm to find a canonical (and hence optimal) DR-plan for all 3 types of 2D qusecs mentioned above (Sections \ref{sec:DRP}, \ref{sec:bodypin}, and \ref{sec:pinnedline}). The canonical DR-plan elucidates the essence of the NP-hardness of finding optimal DR-plans for over-constrained systems. Furthermore, our optimal/canonical DR-plan satisfies desirable properties such as the previously studied Cluster Minimality \cite{hoffman2001decompositionI} (see Figure~\ref{fig:demo_graph:clustmindrp}).

@@

  \item In Section \ref{sec:bodypin} and \ref{sec:pinnedline}, we briefly describe applications of the above techniques to modeling, analyzing, and designing specific properties in 2D material layers~\cite{Jackson2008bodypin}. For Examples \ref{materialexample4} and \ref{materialexample5}, we discuss boundary-conditions for achieving various desired properties of body-hyperpin systems. For Example \ref{materialexample3}, we discuss canonical and optimal DR-plans for pinned line incidence systems \cite{sitharam2014incidence}).

@@

We now briefly survey existing techniques for studying 2D qusecs,  many of which are \dfn{bar-joint} systems (Examples 1 and 2 above, see Sections \ref{sec:prelim}, \ref{sec:DRP}, and \ref{sec:recomb}), \dfn{body-hyperpin} systems (Example 4 and 5, see Section \ref{sec:bodypin}) or \dfn{pinned-line incidence} systems (Example 3, see Section \ref{sec:pinnedline}). The limitations of these techniques directly motivate the contributions of this paper.

@@

\medskip\noindent
\header{(ii) Optimal Recursive Decomposition (DR-planning).}
Recursive decomposition of geometric constraint systems has been formalized \cite{hoffman2001decompositionI,hoffman2001decompositionII} and well-studied \cite{jermann2006decomposition,sitharam2005combinatorial} as the \dfn {Decomposition-Recombination (DR-) planning} problem \seedefsprelim. For the abovementioned classes of 2D qusecs, generic rigidity is a combinatorial property and hence each level of the decomposition should, in principle, be achievable by a graph-based algorithm as in (1), without involving geometric information in the constraint system. Since many such decompositions can exist for a given constraint system, criteria defining desirable or optimal DR-plans and DR-planning algorithms were given in \cite{hoffman2001decompositionI}. An \dfn{optimal DR-plan} is one that minimizes the \dfn{size} \seedefsprelim, i.e., the maximum number of child subsystems of any parent system. Being exponential in the size, the complexity of solving the parent constraint system is overwhelmingly dominated by the complexity of solving the child systems.

@@

\medskip\noindent
\header{(iii) DR-plans for special classes and with other
criteria.}
For a special class of 2D qusecs, namely \dfn{tree-decomposable} systems  \cite{fudos1997graph,owen1991algebraic,joan-arinyo2004revisiting}  common in computer aided mechanical design, (which includes ruler-and-compass and Henneberg-I constructible systems), all DR-plans turn out to be optimal. This satisfies the so-called \dfn{Church-Rosser} criterion, leading to highly efficient DR-planning algorithms. For general 2D qusecs, alternate criteria were suggested such as \dfn{cluster minimality} requiring parent systems to be composed of a minimal set of at least 2 rigid proper subsystems (i.e., no proper subset forms a rigid system); and \dfn{proper maximality}, requiring child subsystems to be maximal rigid proper subsystems of the parent system. \seedefsc.

@@

\medskip\noindent
\header{(iv) Optimal Recombination and Solution Space Navigation.}
For the entire DR-plan, finding all desired solutions is barely tractable even if recombination of solved subsystems is easy for each indecomposable parent system in the DR-plan. This is because even for the simplest, highly decomposable systems with small DR-plans, the problem of finding even a single solution to the input system at the root of the DR-plan is NP-hard  \cite{saxe1979embeddability} and there is a combinatorial explosion of solutions \cite{borcea2004number}. Typically, however, the desired solution has a given orientation type, in which case, the crux of the difficulty is concentrated in the algebraic complexity of (re)combining child system solutions to give a solution to the parent system at any given level of the DR-plan. For fairly general 3D constraint systems, there are algorithms with formal guarantees that uncover underlying matroids to combinatorially obtain an optimal parameterization to minimize the algebraic complexity of the indecomposable parent (sub)systems that occur in the DR-plan \cite{sitharam2010optimized,sitharam2006well,sitharam2010reconciling}, provided the DR-plan meets some of the abovementioned criteria.

@@

\medskip\noindent
\header{(v) Configuration Spaces of Underconstrained Systems.}
For underconstrained 2D bar-joint and body-hyperpin qusecs obtained from various subclasses of tree-decomposable systems, algorithms have been developed to complete them into isostatic systems \cite{joan-arinyo2003transforming,sitharam2005combinatorial,gao2006ctree,sitharam2010convex} and to find paths within the connected components \cite{sitharam2011cayleyI,hidalgo2011reachability} of standard Cartesian configuration spaces. Most of the algorithms with formal guarantees leverage Cayley configuration space theory \cite{sitharam2010convex,sitharam2011cayleyI,sitharam2011cayleyII} to characterize subclasses of graphs and additional constraints that control combinatorial explosion, and provide faithful bijective representation of connected components and paths. These algorithms have decreasing efficiency as the subclass of systems gets bigger, with highest efficiency for underlying partial 2-tree graphs (alternately called tree-width 2, series-parallel, and $K_4$ minor avoiding), moderate efficiency for 1 degree-of-freedom (dof) graphs with low Cayley complexity (which include common linkages such as the Strandbeest, Limacon and Cardioid), and decreased efficiency for general 1-dof tree-decomposable graphs. While software suites exist  \cite{keycurriculum1995geometer,porta2014open,siemens1999d,todd2007geometry}, no such formal algorithms and guarantees are known for non-tree-decomposable systems.

@@

### 2015-03-25 doc/GMP2015/revision_2/s02_preliminaries.tex
Started revision 2 directory. Got rid of \cutout, \addin, and \notetoreviewers macros.

\addin{
Basics definitions and theory for this topic can be found in~\ref{sec:appendix:defs}.

@@

In the remainder of this section and Sections~\ref{sec:DRP} and \ref{sec:recomb} we only consider bar-joint qusecs and graphs. Relevant formal analogies for the other 2 types of qusecs and (hyper)graphs are given in their corresponding sections on applications.
}

@@

\begin{figure*}\centering%
  %
  \begin{subfigure}{0.4\linewidth}\centering
    \includegraphics[height=\myMinHeight]{img/epsfromtikz/demo_graph}
    \caption{}\label{fig:demo_graph:graph}
  \end{subfigure}%
  %
  \hfill
  \begin{subfigure}{0.3\linewidth}\centering
    \includegraphics[height=\myMinHeight]{img/epsfromtikz/demo_graph_comdrp}
    \caption{}\label{fig:demo_graph:comdrp}
  \end{subfigure}%
  %
  \hfill
  \begin{subfigure}{0.3\linewidth}\centering
    \includegraphics[height=\myMinHeight]{img/epsfromtikz/demo_graph_candrp}
    \caption{}\label{fig:demo_graph:candrp}
  \end{subfigure}%
  %
  \caption{\notetoreviewer{This is a new figure.} \addin{(\ref{fig:demo_graph:graph}) A simple graph, $G_{\Delta}$, used to illustrate concepts throughout this and the next section. (\ref{fig:demo_graph:comdrp}) The complete DR-plan of $G_{\Delta}$, i.e.\ $ComDRP(G_{\Delta})$. Dashed lines indicate that the children repeat the same pattern as the others shown on this level. The children of triangles (3 edges) are omitted. (\ref{fig:demo_graph:candrp}) The canonical DR-plan of $G_{\Delta}$, which is optimal (see Section~\ref{sec:DRP}), i.e.\ $OptDRP(G_{\Delta})$. The children of triangles are omitted.}}
  \label{fig:demo_graph}
\end{figure*}%

\begin{figure*}\centering%
  %
  \includegraphics[width=0.4\linewidth]{img/epsfromtikz/demo_graph_clustmindrp}
  \caption{
  \notetoreviewer{This is a new figure.}
  \addin{A DR-plan of $G_{\Delta}$ from Figure~\ref{fig:demo_graph:graph}, which gives cluster minimality. Using the ideas from the paper~\cite{lomonosov2004graph}, the children of a triangle are the primitives of a point and an edge. Thus, the fan-in of this graph is 3, whereas the optimal DR-plans shown in Figure~\ref{fig:demo_graph:candrp} and \ref{fig:demo_graph:candrpseq} have fan-in of 2. With this counterexample, it is clear that cluster minimality is not sufficient for optimality.}
  }
  \label{fig:demo_graph:clustmindrp}
\end{figure*}%

\begin{definition}\label{def:drp}
    The \dfn{decomposition-recombination \mbox{(DR-)} plan} \cite{hoffman2001decompositionI} of graph $G$, $\drp{G}$, \cutout{(see Figure X)} is defined as a forest that has the following properties:
    \begin{enumerate}
        \item Each node represents/contains/is a rigid \cutout{vertex-induced} subgraph of $G$.
        % \item For a node, $C$, that is  a non-trivial graph, its $N$
        % children, $C_1, \ldots, C_N$, are rigid vertex-maximal proper
        % subgraphs of $C$.
        \item Node $C$ has children $C_1,\ldots,C_N$ such that $\bigcup_{i=1}^N{C_i}=C$.
        \item A leaf node is a single vertex. A \dfn{trivial} graph is empty or a single vertex.
        \item A root node is a maximal rigid subgraph of $G$.
    \end{enumerate}
%
    % It can also be described recursively as: the root is $G$, its children are the trivial subgraphs and the DR-plans of its wellconstrained vertex-maximal proper subgraphs whose union is $G$ itself.
%
    A DR-plan is \dfn{complete} if it satisfies an additional rule: for a node $C$ that is  a non-trivial graph, its $N$ children, $C_1, \ldots, C_N$, are all of the rigid vertex-maximal proper subgraphs of $C$. This makes Rule 2 implicit, and in fact, $C$ contains all the edges in the union of its children as well. We denote a complete DR-plan of $G$ as $\comdrp{G}$.

@@

    \addin{See Figures \ref{fig:demo_graph}, \ref{fig:c2c3ofk33s}, \ref{fig:demo_graph:candrpseq}, \ref{fig:demo_graph:clustmindrp}, and \ref{fig:bodypindrp} for examples of DR-plans.}

@@

\cutout{
\noindent
\textbf{Note:} In Section X, $G$ is assumed to be isostatic.
}

@@

\medskip\noindent
\header{Realizing given qusecs.}
Note that this basic task would require finding real solutions \addin{to} a large multivariate polynomial system (of inequalities and equalities representing the constraints). Without DR-planning, this requires double exponential time in the number of variables (even if orientation type is specified). With DR-planning, the complexity is dominated by the size of the largest subsystem that is solved, or recombined from the solutions of its child subsystems\addin{,} i.e., the maximum fan-in occurring in a DR-plan. Thus\addin{,} finding an optimal DR-plan is important for realization.

\medskip\noindent
\header{Recursive block decomposition of the rigidity matrix, stress space and external stresses.}
This can be obtained using the DR-plan\addin{,} \cutout{we obtain a} with each block being the rigidity matrix of an isostatic subgraph. The blocks intersect on isostatic or trivial subgraphs. This recursive block decomposition of the rigidity matrix $R$ correspondingly recursively decomposes the stress vectors $s$ that balance a given external stress $t$ since $sR = t$. Moreover, the external stress vector $t$ is recursively ``distributed'' into external stresses for each node of the DR-plan. When the rigidity matrix and stress vectors contain indeterminates, this  can be used to design underlying graph and parameters $\delta$ to obtain an optimal distribution of stresses on the geometric primitives. When the realization of qusecs is given, this can be used to analyze the distribution of a given external load.

>>>>

Basics definitions and theory for this topic can be found in~\ref{sec:appendix:defs}.

@@

In the remainder of this section and Sections~\ref{sec:DRP} and \ref{sec:recomb} we only consider bar-joint qusecs and graphs. Relevant formal analogies for the other 2 types of qusecs and (hyper)graphs are given in their corresponding sections on applications.

@@

\begin{figure*}\centering%
  %
  \begin{subfigure}{0.4\linewidth}\centering
    \includegraphics[height=\myMinHeight]{img/epsfromtikz/demo_graph}
    \caption{}\label{fig:demo_graph:graph}
  \end{subfigure}%
  %
  \hfill
  \begin{subfigure}{0.3\linewidth}\centering
    \includegraphics[height=\myMinHeight]{img/epsfromtikz/demo_graph_comdrp}
    \caption{}\label{fig:demo_graph:comdrp}
  \end{subfigure}%
  %
  \hfill
  \begin{subfigure}{0.3\linewidth}\centering
    \includegraphics[height=\myMinHeight]{img/epsfromtikz/demo_graph_candrp}
    \caption{}\label{fig:demo_graph:candrp}
  \end{subfigure}%
  %
  \caption{(\ref{fig:demo_graph:graph}) A simple graph, $G_{\Delta}$, used to illustrate concepts throughout this and the next section. (\ref{fig:demo_graph:comdrp}) The complete DR-plan of $G_{\Delta}$, i.e.\ $ComDRP(G_{\Delta})$. Dashed lines indicate that the children repeat the same pattern as the others shown on this level. The children of triangles (3 edges) are omitted. (\ref{fig:demo_graph:candrp}) The canonical DR-plan of $G_{\Delta}$, which is optimal (see Section~\ref{sec:DRP}), i.e.\ $OptDRP(G_{\Delta})$. The children of triangles are omitted.}
  \label{fig:demo_graph}
\end{figure*}%

\begin{figure*}\centering%
  %
  \includegraphics[width=0.4\linewidth]{img/epsfromtikz/demo_graph_clustmindrp}
  \caption{A DR-plan of $G_{\Delta}$ from Figure~\ref{fig:demo_graph:graph}, which gives cluster minimality. Using the ideas from the paper~\cite{lomonosov2004graph}, the children of a triangle are the primitives of a point and an edge. Thus, the fan-in of this graph is 3, whereas the optimal DR-plans shown in Figure~\ref{fig:demo_graph:candrp} and \ref{fig:demo_graph:candrpseq} have fan-in of 2. With this counterexample, it is clear that cluster minimality is not sufficient for optimality.}
  \label{fig:demo_graph:clustmindrp}
\end{figure*}%

\begin{definition}\label{def:drp}
    The \dfn{decomposition-recombination \mbox{(DR-)} plan} \cite{hoffman2001decompositionI} of graph $G$, $\drp{G}$, is defined as a forest that has the following properties:
    \begin{enumerate}
        \item Each node represents/contains/is a rigid subgraph of $G$.
        % \item For a node, $C$, that is  a non-trivial graph, its $N$
        % children, $C_1, \ldots, C_N$, are rigid vertex-maximal proper
        % subgraphs of $C$.
        \item Node $C$ has children $C_1,\ldots,C_N$ such that $\bigcup_{i=1}^N{C_i}=C$.
        \item A leaf node is a single vertex. A \dfn{trivial} graph is empty or a single vertex.
        \item A root node is a maximal rigid subgraph of $G$.
    \end{enumerate}
%
    % It can also be described recursively as: the root is $G$, its children are the trivial subgraphs and the DR-plans of its wellconstrained vertex-maximal proper subgraphs whose union is $G$ itself.
%
    A DR-plan is \dfn{complete} if it satisfies an additional rule: for a node $C$ that is  a non-trivial graph, its $N$ children, $C_1, \ldots, C_N$, are all of the rigid vertex-maximal proper subgraphs of $C$. This makes Rule 2 implicit, and in fact, $C$ contains all the edges in the union of its children as well. We denote a complete DR-plan of $G$ as $\comdrp{G}$.

@@

    See Figures \ref{fig:demo_graph}, \ref{fig:c2c3ofk33s}, \ref{fig:demo_graph:candrpseq}, \ref{fig:demo_graph:clustmindrp}, and \ref{fig:bodypindrp} for examples of DR-plans.

@@

@@

\medskip\noindent
\header{Realizing given qusecs.}
Note that this basic task would require finding real solutions to a large multivariate polynomial system (of inequalities and equalities representing the constraints). Without DR-planning, this requires double exponential time in the number of variables (even if orientation type is specified). With DR-planning, the complexity is dominated by the size of the largest subsystem that is solved, or recombined from the solutions of its child subsystems, i.e., the maximum fan-in occurring in a DR-plan. Thus, finding an optimal DR-plan is important for realization.

\medskip\noindent
\header{Recursive block decomposition of the rigidity matrix, stress space and external stresses.}
This can be obtained using the DR-plan, with each block being the rigidity matrix of an isostatic subgraph. The blocks intersect on isostatic or trivial subgraphs. This recursive block decomposition of the rigidity matrix $R$ correspondingly recursively decomposes the stress vectors $s$ that balance a given external stress $t$ since $sR = t$. Moreover, the external stress vector $t$ is recursively ``distributed'' into external stresses for each node of the DR-plan. When the rigidity matrix and stress vectors contain indeterminates, this  can be used to design underlying graph and parameters $\delta$ to obtain an optimal distribution of stresses on the geometric primitives. When the realization of qusecs is given, this can be used to analyze the distribution of a given external load.
